---
title: "02_dplyr"
author: "Pao"
date: "July 27, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Notas del capítulo 5

Data transformation with dpylr

```{r}
library(nycflights13)
library(tidyverse)
```

```{r}
flights
```

### Verbos de dyplr

* `filter()` filtran las filas que cumplen la condición o condiciones dadas.
* `arrange()` ordena las filas en orden ascendente o descendente (si agregamos `desc()`) según la columna o columnas indicadas.
* `select()` permite seleccionar variables dentro de un dataframe. Acepta funciones para seleccionar los nombres de las variables como `starts_with("abc")`. El hermano es `rename()` que se usa para renombrar las variables.
* `mutate()` permite agregar nuevas columnas. Si solo querés quedarte con las nuevas usas `transmute()`.
* `summarise()` colapse el data frame a una fila usando alguna función. Lo interesante es usarlo junto con `group_by()`.

### Ejercicios

1. Find all flights that

    1. Had an arrival delay of two or more hours
    2. Flew to Houston (IAH or HOU)
    3. Were operated by United, American, or Delta
    4. Departed in summer (July, August, and September)
    5. Arrived more than two hours late, but didn’t leave late
    6. Were delayed by at least an hour, but made up over 30 minutes in flight
    7. Departed between midnight and 6am (inclusive)

```{r}
filter(flights, arr_delay >= 120)

filter(flights, dest %in% c("IAH", "HOU"))

filter(flights, carrier %in% c("UA", "DL"))

filter(flights, month %in% c(7, 8, 9))

filter(flights, arr_delay >= 120 & dep_delay == 0)
```

2. Another useful dplyr filtering helper is `between()`. What does it do? Can you use it to simplify the code needed to answer the previous challenges?

```{r}
filter(flights, between(month, 7, 9))
```

3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
count(filter(flights, is.na(dep_time)))

filter(flights, is.na(dep_time))
```

4. Why is `NA ^ 0` not missing? Why is `NA | TRUE` not missing? Why is `FALSE & NA` not missing? Can you figure out the general rule? (`NA * 0` is a tricky counterexample!)

```{r}
NA^0
NA | TRUE      # NA o TRUE, es TRUE
FALSE & NA     # Cualquier cosa y falso es FALSE

```

1. How could you use `arrange()` to sort all missing values to the start? (Hint: use is.na()).

```{r}
arrange(flights, desc(dep_time)) # No funciona

arrange(flights, desc(is.na(dep_time)), dep_time) # TRUE > FALSE
```

2. Sort flights to find the most delayed flights. Find the flights that left earliest.

```{r}
arrange(flights, desc(dep_delay))
```

3. Sort flights to find the fastest flights.

```{r}
fastest_flights <- mutate(flights, mph = distance / air_time * 60)
fastest_flights <- select(
  fastest_flights, mph, distance, air_time,
  flight, origin, dest, year, month, day
)

arrange(fastest_flights, desc(mph))
```

1. Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.

```{r}
# Mencionando uno a uno, fiaca.
# Expresión regular 3ch_ch?, en realidad funciona: "^(dep|arr)_(time|delay)$"
select(flights, matches("^(dep|arr)_(time|delay)$"))
```

2. What happens if you include the name of a variable multiple times in a select() call?

```{r}
select(flights, air_time, air_time)
```

4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

```{r}
select(flights, contains("TIME")) # Por defecto contains() ignora mayuscualas/minúsculas
```

1. Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}
mutate(flights, 
       dep_hour = dep_time %/% 100,
       dep_min = dep_time %% 100,
       dep_min_00 = (dep_hour*60 + dep_min) %% 1440) 
```

2. Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?

```{r}
mutate(flights,
       dep_time = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
       arr_time = (arr_time %/% 100 * 60 + arr_time %% 100) %% 1440,
       air_time2 = arr_time - dep_time) %>% 
  select(air_time, air_time2)
``` 

