---
title: "03_tibbles"
author: "Pao"
date: "8/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Notas del capítulo 5

Conociendo las tibbles. Funcionan como data.frames pero son más "seguras". Por ejemplo si haces el subset de una columna con `[]` un data.frame te devuelve un vector pero una tiblle te devuelve una tibble. También te permite poner nombres raros a las columnas, no te convierte strings a factor, y no hace *partial maching*.

### Ejercicios

1. How can you tell if an object is a tibble?

```{r}
class(mtcars)
is_tibble(mtcars)
```

2. Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?

```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]

```
Esto tiene que ver con el *macheo parcial*, cuando le pedimos que nos muestre la columna x, no la encuentra pero encuentra una parecida y entonces asume que es esa. Esto puede ser muy práctico si querés ahorrarte un par de letras pero si no era tu intención, you are in trouble. 


3. If you have the name of a variable stored in an object, e.g. `var <- "mpg"`, how can you extract the reference variable from a tibble?

```{r}
var <- "mpg"

mtcars <- as_tibble(mtcars)
mtcars[[var]]

```


4. Practice referring to non-syntactic names in the following data frame by:

    - Extracting the variable called 1.

    - Plotting a scatterplot of 1 vs 2.

    - Creating a new column called 3 which is 2 divided by 1.

    - Renaming the columns to one, two and three.

```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)

annoying[["1"]]
annoying$`1`

ggplot(annoying, aes(`1`, `2`)) +
  geom_point()

annoying %>% 
  mutate(`3` = `2`/`1`)

annoying <- rename(annoying, one = `1`, two = `2`, three = `3`)

```

5. What does `tibble::enframe()` do? When might you use it?

enframe() converts named atomic vectors or lists to one- or two-column data frames. For a list, the result will be a nested tibble with a column of type list. For unnamed vectors, the natural sequence is used as name column.

```{r}
tibble::enframe(c(a = 1, b = 2, c = 3))
```


6. What option controls how many additional column names are printed at the footer of a tibble?

The help page for the print() method of tibble objects is discussed in ?print.tbl. The n_extra argument determines the number of extra columns to print information for.